---
output: 
  html_document: 
    fig_width: 9
    fig_height: 5
    keep_md: yes
---

```{r Set wd load and clean data, warning=FALSE, , include=FALSE}
#Note to run the animation you will have to instal Imagemagik from http://www.imagemagick.org/script/binary-releases.php
# when you install Imagemagik remember to select the "instal legacy files" option so convert.exe is included

#Set working directory where this script and the raw excel file are saved 
setwd("C:/Users/Dan/Documents/GitHub/MelbournePropertyPrices")

### Install/load required packages
#List of R packages required for this analysis:
required_packages <- c("ggmap", "readr", "ggplot2", "dplyr", "readxl","tidyr",
                       "stringr", "gganimate","animation","ggiraph", "viridis")
#Install required_packages:
new.packages <- required_packages[!(required_packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
#Load required_packages:
lapply(required_packages, require, character.only = TRUE)

#Set decimal points and disable scientific notation
options(digits=8, scipen=999) 

###Load Map
# Can download using get_map function:
# Melbourne <- get_map(location = 'Melbourne, Australia', zoom = 10, maptype="terrain")
# But I saved it to workind directory so can load it and work offline:
load("Melbourne.rda")

###Load Apartment price data
# Got this data here http://www.dtpli.vic.gov.au/property-and-land-titles/property-information/property-prices ...
# ...but saved it to workind directory so can load it and work offline
suburb_Apartment_2015<-read_excel("suburb_unit_2015.xls", col_names=F, skip = 3) %>% 
            na.omit() %>%
            distinct() %>%
            mutate(`Property Type`="Apartment")
#Load House Price Data and row bind it to Apartment price data 
property_2005_2015<-read_excel("suburb_house_2015.xls", col_names=F, skip = 3) %>% 
            na.omit() %>%
            distinct() %>%
            mutate(`Property Type`="House") %>%
            bind_rows(suburb_Apartment_2015)
rm(suburb_Apartment_2015)


#rename the columns 
names(property_2005_2015)<-c("suburb", 2005:2015, "Prelim 2016", "Change 2014-2015 (%)", "Change 2005-2015 (%)", "Growth (%) PA 2005-2015","Property Type")

# assuming that "-" and "0" means data not available (NA), 
property_2005_2015[property_2005_2015 == 0] <- NA
property_2005_2015[property_2005_2015 == "-"] <- NA


## Read in the lat/long data:
# Also saved it to workind directory so can load it and work offline:
lat_long<-read_csv("Australian_Post_Codes_Lat_Lon.csv")  %>% 
    mutate(postcode=as.character(postcode)) %>%
    distinct() %>%
    select(-dc, -type)


#create VIC only lat_long
VIC_lat_long<- lat_long %>% 
    filter(state=="VIC") %>%
    select(-postcode) %>%
    distinct() %>% 
    filter(suburb %in%  property_2005_2015$suburb)

#Merge VIC_lat_long into property_2005_2015
property_2005_2015<- full_join(property_2005_2015, VIC_lat_long,  by=c("suburb"))
rm(VIC_lat_long)


property_2005_2015 <- property_2005_2015 %>%
    mutate(`Change 2014-2015 (%)` = as.numeric(`Change 2014-2015 (%)`),
           `Change 2005-2015 (%)` = as.numeric(`Change 2005-2015 (%)`),
           `Growth (%) PA 2005-2015` = as.numeric(`Growth (%) PA 2005-2015`),
           `Change 2005-2015 ($)` = as.numeric(`2015`)-as.numeric(`2005`)) %>%
     distinct()
    

#melt from wide to long format
property_2005_2015<-property_2005_2015 %>% 
    gather(key=Year, value=`Median Price ($)`, -suburb, -lon, -lat, -state, -`Property Type`, -`Change 2014-2015 (%)`, -`Change 2005-2015 (%)`, -`Growth (%) PA 2005-2015`, -`Change 2005-2015 ($)`) %>%
    mutate(`Median Price ($)`=as.numeric(`Median Price ($)`)) %>%
    distinct()





```




```{r map, warning=FALSE, include=FALSE}

p1<-ggmap(Melbourne) +
    geom_point(data = property_2005_2015,
               aes(x =lon, y= lat, frame = Year, size=`Median Price ($)`,
                   colour = `Median Price ($)`), alpha=.75, shape="$") +
        # scale_colour_gradientn(colours=rainbow(5)) +
    scale_color_viridis() +
        scale_radius (range = c(4, 14), trans = "identity", guide = "legend") +
    facet_wrap(~`Property Type`) +
    ggtitle("Median Melbourne Property Prices ($) from 2005-2016 \n")

p1 <- p1 + theme(aspect.ratio=1) +
        theme(axis.title.x = element_blank(),
            axis.text.x  =  element_blank(),
            axis.title.x = element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.y  =  element_blank(),
            axis.title.y = element_blank(),
            axis.ticks.y=element_blank(),
            legend.title = element_text(size=12, face="bold"),
            legend.text = element_text(size = 12, face = "bold"),
            strip.text.x = element_text(size=12, face="bold"),
            plot.title = element_text(size = 14, face = "bold"),
            legend.position="right")

#Save the .gif to working directory
gg_animate(p1, 'Melbourne.gif', ani.width = 1000, ani.height = 600, interval = 0.75)

#show the gif in the graphics pannel
gg_animate(p1)


```


##-Scroll your mouse to zoom in and out on the maps
##-Hover over a point to see more information about that suburb
##-Click on a suburb to go to a new google maps window/tab

```{r still Map, warning=FALSE, echo=FALSE}
#####create column with text to appear with 'tooltip' mouse hover###
#Growth (%) PA 2005-2015`
property_2005_2015$tooltip_point1 <- paste0(
    "<b>", property_2005_2015$suburb, "</b>",
    "<br>Growth per annum:",
    "<br>",property_2005_2015$`Growth (%) PA 2005-2015`,"%",
    "</span></div>")   
#`Change 2005-2015 (%)`
property_2005_2015$tooltip_point2 <- paste0(
    "<b>", property_2005_2015$suburb, "</b>",
    "<br>Change in Median Price:",
    "<br>",property_2005_2015$`Change 2005-2015 (%)`,"%",
    "</span></div>")   

property_2005_2015$tooltip_point3 <- paste0(
    "<b>", property_2005_2015$suburb, "</b>",
    "<br>Change in Median Price:",
    "<br>$",property_2005_2015$`Change 2005-2015 ($)`,
    "</span></div>")   


# create an 'onclick' column
property_2005_2015<-property_2005_2015 %>% mutate(google_map_lookup = paste(suburb, state, "Australia", sep=", "))
property_2005_2015$onclick <- sprintf("window.open(\"%s%s\")",
  "https://www.google.com.au/maps/place/", as.character(property_2005_2015$google_map_lookup))



#Make plots
p2<-ggmap(Melbourne) + 
    geom_point_interactive(data = property_2005_2015,
                           aes(x =lon, y= lat,
                               size= `Growth (%) PA 2005-2015`, 
                               colour = `Growth (%) PA 2005-2015`, 
                               tooltip=tooltip_point1, 
                               data_id = tooltip_point1,
                               onclick = onclick)) +
    scale_color_viridis() +
    scale_radius (range = c(1, 5), trans = "identity", guide = "legend") +
    facet_wrap(~`Property Type`) +
    ggtitle("Growth per annum in Median Melbourne \n Property Prices from 2005-2015 \n") +
    theme(axis.title.x = element_blank(), 
        axis.text.x  =  element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y  =  element_blank(), 
        axis.title.y = element_blank(),
        axis.ticks.y=element_blank(),
        legend.title = element_text(face="bold"),
        legend.text = element_text(face = "bold"),
        strip.text.x = element_text(face="bold"),
        plot.title = element_text(face = "bold"))  
ggiraph(code = {print(p2)}, 
        zoom_max = 5,
        width_svg = 8, height_svg = 6,
        tooltip_offx = 20, tooltip_offy = -10, 
        hover_css = "fill:black;",
        tooltip_opacity = 0.7)



p3<-ggmap(Melbourne) + 
    geom_point_interactive(data = property_2005_2015,
                           aes(x =lon, y= lat,
                               size= `Change 2005-2015 (%)`, 
                               colour = `Change 2005-2015 (%)`, 
                               tooltip=tooltip_point2, 
                               data_id = tooltip_point2,
                               onclick = onclick)) +
    scale_color_viridis() +
    scale_radius (range = c(1, 6), trans = "identity", guide = "legend") +
    facet_wrap(~`Property Type`) +
    ggtitle("Total change (%) in Median Melbourne \n Property Prices from 2005-2015 \n") +
    theme(axis.title.x = element_blank(), 
        axis.text.x  =  element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y  =  element_blank(), 
        axis.title.y = element_blank(),
        axis.ticks.y=element_blank(),
        legend.title = element_text(face="bold"),
        legend.text = element_text(face = "bold"),
        strip.text.x = element_text(face="bold"),
        plot.title = element_text(face = "bold"))  
ggiraph(code = {print(p3)}, 
        zoom_max = 5,
        width_svg = 8, height_svg = 6,
        tooltip_offx = 20, tooltip_offy = -10, 
        hover_css = "fill:black;",
        tooltip_opacity = 0.7)



p4<-ggmap(Melbourne) + 
    geom_point_interactive(data=property_2005_2015, 
                                aes(x =lon, y= lat,
                                size= `Change 2005-2015 ($)`, 
                                colour = `Change 2005-2015 ($)`, 
                                tooltip=tooltip_point3, 
                                data_id = tooltip_point3,
                                onclick = onclick)) +
    scale_color_viridis() +
    scale_radius (range = c(2, 8), trans = "identity", guide = "legend") +
    facet_wrap(~`Property Type`) +
    ggtitle("Total change ($) in Median Melbourne \n Property Prices from 2005-2015 \n") +
    theme(axis.title.x = element_blank(), 
        axis.text.x  =  element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y  =  element_blank(), 
        axis.title.y = element_blank(),
        axis.ticks.y=element_blank(),
        legend.title = element_text(face="bold"),
        legend.text = element_text(face = "bold"),
        strip.text.x = element_text(face="bold"),
        plot.title = element_text(face = "bold"))  
ggiraph(code = {print(p4)}, 
        zoom_max = 5,
        width_svg = 8, height_svg = 6,
        tooltip_offx = 20, tooltip_offy = -10, 
        hover_css = "fill:black;",
        tooltip_opacity = 0.7)



  
```


##Static versions of the interactive maps above

Notice the background of the static versions render properly.
Hopfully I'll get the backgrounds rendering properly on the interactive plots above an update this blog post with it very soon!
```{r interactive Map, warning=FALSE, echo=FALSE}

ggmap(Melbourne) + 
    geom_point(data = property_2005_2015, 
               aes(x =lon, y= lat, 
                   size=`Growth (%) PA 2005-2015`, 
                   colour = `Growth (%) PA 2005-2015`), 
               alpha=.75) +
    scale_color_viridis() +
    scale_radius (range = c(1, 5), trans = "identity", guide = "legend") +
    facet_wrap(~`Property Type`) +
    ggtitle("Growth per annum (%) in Median Melbourne \n Property Prices from 2005-2015 \n") +
    theme(axis.title.x = element_blank(), 
        axis.text.x  =  element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y  =  element_blank(), 
        axis.title.y = element_blank(),
        axis.ticks.y=element_blank(),
        legend.title = element_text(face="bold"),
        legend.text = element_text(face = "bold"),
        strip.text.x = element_text(face="bold"),
        plot.title = element_text(face = "bold"))


ggmap(Melbourne) + 
    geom_point(data = property_2005_2015, 
               aes(x =lon, y= lat,
                   size=`Change 2005-2015 (%)`, 
                   colour = `Change 2005-2015 (%)`), 
               alpha=.75) +
    scale_color_viridis() +
    scale_radius (range = c(1, 6), trans = "identity", guide = "legend") +
    facet_wrap(~`Property Type`) +
    ggtitle("Total change (%) in Median Melbourne \n Property Prices from 2005-2015 \n") +
    theme(axis.title.x = element_blank(), 
        axis.text.x  =  element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y  =  element_blank(), 
        axis.title.y = element_blank(),
        axis.ticks.y=element_blank(),
        legend.title = element_text(face="bold"),
        legend.text = element_text(face = "bold"),
        strip.text.x = element_text(face="bold"),
        plot.title = element_text(face = "bold"))



ggmap(Melbourne) + 
    geom_point(data = property_2005_2015, 
               aes(x =lon, y= lat,
                   size=`Change 2005-2015 ($)`, 
                   colour = `Change 2005-2015 ($)`), 
               alpha=.75) +
    scale_color_viridis() +
    scale_radius (range = c(2, 8), trans = "identity", guide = "legend") +
    facet_wrap(~`Property Type`) +
    ggtitle("Total change ($) in Median Melbourne \n Property Prices from 2005-2015 \n") +
    theme(axis.title.x = element_blank(), 
        axis.text.x  =  element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x=element_blank(),
        axis.text.y  =  element_blank(), 
        axis.title.y = element_blank(),
        axis.ticks.y=element_blank(),
        legend.title = element_text(face="bold"),
        legend.text = element_text(face = "bold"),
        strip.text.x = element_text(face="bold"),
        plot.title = element_text(face = "bold"))

```

